### 2017-12-11 submission/manuscript.Rmd
change paragraph order
remove/replace role with associations where appropriate

**Conclusions.** This data provides support for the importance of the bacteral community to both adenoma and carcinoma genesis. The evidence collected within this study on the role of the microbiota in CRC pathogenesis is much stronger for carcinoma then adenoma. A strong reason for this may be in part due to the low power to detect more subtle changes in the majority of studies that have been performed to date.

@@

Colorectal cancer (CRC) is a growing world-wide health problem [@siegel_cancer_2016] in which the microbiota has been purported to play an active role in disease pathogenesis [@flynn_metabolic_2016]. Numerous studies have shown the importance of both individual microbes [@goodwin_polyamine_2011; @abed_fap2_2016; @arthur_intestinal_2012; @kostic_fusobacterium_2013; @wu_human_2009] and the overall community [@zackular_manipulation_2016; @zackular_gut_2013; @baxter_structure_2014] in polyp formation using mouse models of CRC. There have also been numerous case/control studies investigating the microbiota in the formation of both adenoma and carcinoma. Recently, a meta-analysis was published investigating whether specific biomarkers could be consistently identified using multiple data sets [@shah_leveraging_2017]. Many of the studies, along with the previous meta-analysis, focus on identifying biomarkers or individual microbes but do not critically investigate the role the community has in CRC.

@@

Although there are still questions that need to be answered regarding the microbiota and carcinoma, a clearer framework for their relationship is beginning to develop as to how this occurs. From our observations many changes in carcinoma could easily result in effect sizes that are 30% or more between the case and control and most studies analyzed have sufficient power to detect these types of changes [Figure 7]. Conversely, the role of the microbiota in adenoma is less clear and part of the reason this may be is because many studies are not powered effectively to observe the small changes reported here. None of the studies analyzed were properly powered to detect a 10% or lower change between case and controls. This small effect size range may well be the scope in which differences consistently occur in adenoma due to the subtle changes in community that occur between control and adenoma. Future studies investigating adenoma and the microbiota need to take these factors into consideration if we are to work out how the microbiota contributes to adenoma formation.

>>>>

**Conclusions.** This data provides support for the importance of the bacteral community to both adenoma and carcinoma genesis. The evidence collected within this study on the associations between the microbiota and CRC are much stronger for carcinoma then adenoma. A strong reason for this may be in part due to the low power to detect more subtle changes in the majority of studies that have been performed to date.

@@

Colorectal cancer (CRC) is a growing world-wide health problem [@siegel_cancer_2016] in which the microbiota has been purported to play an active role in disease pathogenesis [@flynn_metabolic_2016]. Numerous studies have shown the importance of both individual microbes [@goodwin_polyamine_2011; @abed_fap2_2016; @arthur_intestinal_2012; @kostic_fusobacterium_2013; @wu_human_2009] and the overall community [@zackular_manipulation_2016; @zackular_gut_2013; @baxter_structure_2014] in polyp formation using mouse models of CRC. There have also been numerous case/control studies investigating the microbiota in the formation of both adenoma and carcinoma. Recently, a meta-analysis was published investigating whether specific biomarkers could be consistently identified using multiple data sets [@shah_leveraging_2017]. Many of the studies, along with the previous meta-analysis, focus on identifying biomarkers or individual microbes but do not critically investigate how the community changes in CRC.

@@

Although there are still questions that need to be answered regarding the microbiota and carcinoma, a clearer framework for their relationship is beginning to develop as to how this occurs. From our observations many changes in carcinoma could easily result in effect sizes that are 30% or more between the case and control and most studies analyzed have sufficient power to detect these types of changes [Figure 7]. Conversely, the associations between the microbiota and adenoma is less clear and part of the reason this may be is because many studies are not powered effectively to observe the small changes reported here. None of the studies analyzed were properly powered to detect a 10% or lower change between case and controls. This small effect size range may well be the scope in which differences consistently occur in adenoma due to the subtle changes in community that occur between control and adenoma. Future studies investigating adenoma and the microbiota need to take these factors into consideration if we are to work out how the microbiota contributes to adenoma formation.

### 2017-12-11 submission/manuscript.tex
change paragraph order
remove/replace role with associations where appropriate

\textbf{Conclusions.} This data provides support for the importance of
the bacteral community to both adenoma and carcinoma genesis. The
evidence collected within this study on the role of the microbiota in
CRC pathogenesis is much stronger for carcinoma then adenoma. A strong
reason for this may be in part due to the low power to detect more
subtle changes in the majority of studies that have been performed to
date.

@@

Colorectal cancer (CRC) is a growing world-wide health problem {[}1{]}
in which the microbiota has been purported to play an active role in
disease pathogenesis {[}2{]}. Numerous studies have shown the importance
of both individual microbes {[}3--7{]} and the overall community
{[}8--10{]} in polyp formation using mouse models of CRC. There have
also been numerous case/control studies investigating the microbiota in
the formation of both adenoma and carcinoma. Recently, a meta-analysis
was published investigating whether specific biomarkers could be
consistently identified using multiple data sets {[}11{]}. Many of the
studies, along with the previous meta-analysis, focus on identifying
biomarkers or individual microbes but do not critically investigate the
role the community has in CRC.

@@

Although there are still questions that need to be answered regarding
the microbiota and carcinoma, a clearer framework for their relationship
is beginning to develop as to how this occurs. From our observations
many changes in carcinoma could easily result in effect sizes that are
30\% or more between the case and control and most studies analyzed have
sufficient power to detect these types of changes {[}Figure 7{]}.
Conversely, the role of the microbiota in adenoma is less clear and part
of the reason this may be is because many studies are not powered
effectively to observe the small changes reported here. None of the
studies analyzed were properly powered to detect a 10\% or lower change
between case and controls. This small effect size range may well be the
scope in which differences consistently occur in adenoma due to the
subtle changes in community that occur between control and adenoma.
Future studies investigating adenoma and the microbiota need to take
these factors into consideration if we are to work out how the
microbiota contributes to adenoma formation.

>>>>

\textbf{Conclusions.} This data provides support for the importance of
the bacteral community to both adenoma and carcinoma genesis. The
evidence collected within this study on the associations between the
microbiota and CRC are much stronger for carcinoma then adenoma. A
strong reason for this may be in part due to the low power to detect
more subtle changes in the majority of studies that have been performed
to date.

@@

Colorectal cancer (CRC) is a growing world-wide health problem {[}1{]}
in which the microbiota has been purported to play an active role in
disease pathogenesis {[}2{]}. Numerous studies have shown the importance
of both individual microbes {[}3--7{]} and the overall community
{[}8--10{]} in polyp formation using mouse models of CRC. There have
also been numerous case/control studies investigating the microbiota in
the formation of both adenoma and carcinoma. Recently, a meta-analysis
was published investigating whether specific biomarkers could be
consistently identified using multiple data sets {[}11{]}. Many of the
studies, along with the previous meta-analysis, focus on identifying
biomarkers or individual microbes but do not critically investigate how
the community changes in CRC.

@@

Although there are still questions that need to be answered regarding
the microbiota and carcinoma, a clearer framework for their relationship
is beginning to develop as to how this occurs. From our observations
many changes in carcinoma could easily result in effect sizes that are
30\% or more between the case and control and most studies analyzed have
sufficient power to detect these types of changes {[}Figure 7{]}.
Conversely, the associations between the microbiota and adenoma is less
clear and part of the reason this may be is because many studies are not
powered effectively to observe the small changes reported here. None of
the studies analyzed were properly powered to detect a 10\% or lower
change between case and controls. This small effect size range may well
be the scope in which differences consistently occur in adenoma due to
the subtle changes in community that occur between control and adenoma.
Future studies investigating adenoma and the microbiota need to take
these factors into consideration if we are to work out how the
microbiota contributes to adenoma formation.
